---
title: "Thesis code"
author: "Hassan Aziz"
date: "2023-10-29"
output: html_document
---

## Loading Packages
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(gplots)
library(dplyr)
library(arsenal)
library(knitr) 
library(stargazer)
library(effects)
library(readxl)
library(reshape2)
library(plm)
library(glmnet)
library(scales)
library(lmtest)
library(urca)
library(forecast)
library(tseries)
library(systemfit)
library(writexl)
library(janitor)
library(sandwich)
library(broom)
library(WDI)
library(table1)
library(printr)
options(scipen = 999)  
```

## WDI data
### Load the dataset
```{r}
wdi <- read_excel("P_Data_Extract_From_World_Development_Indicators-9.xlsx")

```
### Clean the dataset
```{r}
### Rename variables
names(wdi)[names(wdi) == "Country Name"] <- "country"
names(wdi)[names(wdi) == "Country Code"] <- "country_code"
names(wdi)[names(wdi) == "Series Name"] <- "series"

# Drop irrelevant variables
wdi$`Series Code` <- NULL

# Remove everything after the first space from column names using sub()
colnames(wdi) <- sub(" .*", "", colnames(wdi))

# Define the columns you want to exclude from conversion
columns_to_exclude <- c("country", "country_code", "series")

# Get the names of columns to convert (all except those in columns_to_exclude)
columns_to_convert <- setdiff(names(wdi), columns_to_exclude)

# Convert all columns except the specified ones to numeric
wdi[columns_to_convert] <- lapply(wdi[columns_to_convert], as.numeric)

# Change data from wide to long
wdi_long <- wdi %>%
  pivot_longer(cols = -c(country, country_code, series), 
               names_to = "Year",
               values_to = "Value")

# remove missing values
wdi_long <- na.omit(wdi_long)

# Change data from long back to wide
wdi_wide <- wdi_long %>%
  spread(key = series, value = Value)

names(wdi_wide)[names(wdi_wide) == "Year"] <- "year"
wdi_wide$year <- as.numeric(wdi_wide$year)
```

## Our World Data
### Loading and cleaning the dataset
```{r}
primary_energy <-read.csv("primary-energy-cons.csv")
emissions <- read.csv("annual-co2-emissions-per-country.csv")
temp <- merge(primary_energy, emissions, by = c("Entity", "Code", "Year"), all = FALSE)
names(temp)[names(temp) == "Primary.energy.consumption..TWh."] <- "energy_consumption(TWh)"
names(temp)[names(temp) == "Annual.CO..emissions"] <- "co2_emissions_our_world"
names(temp)[names(temp) == "Entity"] <- "country"
names(temp)[names(temp) == "Code"] <- "country_code"
names(temp)[names(temp) == "Year"] <- "year"
```
### Merging the two datasets
```{r}
co2 <- merge(wdi_wide, temp, by = c("country", "country_code", "year"), all = FALSE)
co2$country_code <- NULL
```

### Scaling the data
```{r}
co2$year <- as.numeric(co2$year)
## Log transformations
co2$lnCO2_wdi <- log(co2$`CO2 emissions (kt)`)
co2$lnCO2 <- log(co2$co2_emissions_our_world)

co2$ln_forest <- log(co2$`Forest area (sq. km)`)
co2$ln_agri <- log(co2$`Cereal production (metric tons)`)
co2$lnGDP <- log(co2$`GDP (constant 2015 US$)`)
co2$ln_population <- log(co2$`Population, total`)
co2$ln_urban <- log(co2$`Urban population (% of total population)`)
co2$ln_energy <- log(co2$`energy_consumption(TWh)`)
co2$gross_cap_by_gdp <- ((co2$`Gross fixed capital formation (constant 2015 US$)`)/(co2$`GDP (constant 2015 US$)`))*100
co2$ln_capital <- log(co2$gross_cap_by_gdp)
co2$ln_fossil <- log(co2$`Fossil fuel energy consumption (% of total)`)
co2$lnGDPsq <- log((co2$`GDP (constant 2015 US$)`)^2)
```

## Cleaning and tests
### Creating four datasets
```{r}
co2 <- co2[, !names(co2) %in% "Foreign direct investment, net (BoP, current US$)"]
# 
# # Without Nepal but with all factors including GFCF
# sann <- co2[co2$country != "Nepal", ]
# sann <- na.omit(sann)

# All South Asia but without GFCF
sanog <- co2[, !names(co2) %in% "gross_cap_by_gdp"]
sanog <- sanog[, !names(sanog) %in% "ln_capital"]
sanog <- sanog[, !names(sanog) %in% "Gross fixed capital formation (constant 2015 US$)"]
sanog <- na.omit(sanog)

# # All South Asia but without GFCF and Fossil Fuel
# sanf <- co2[, !names(co2) %in% "gross_cap_by_gdp"]
# sanf <- sanf[, !names(sanf) %in% "ln_capital"]
# sanf <- sanf[, !names(sanf) %in% "Gross fixed capital formation (constant 2015 US$)"]
# sanf <- sanf[, !names(sanf) %in% "Fossil fuel energy consumption (% of total)"]
# sanf <- sanf[, !names(sanf) %in% "ln_fossil"]
# sanf <- na.omit(sanf)
# 
# # All factors except Fossil fuel, no nepal
# sann_nof <- co2[co2$country != "Nepal", ]
# sann_nof <- sann_nof[, !names(sann_nof) %in% "Fossil fuel energy consumption (% of total)"]
# sann_nof <- sann_nof[, !names(sann_nof) %in% "ln_fossil"]
# sann_nof <- na.omit(sann_nof)
```

### Creating panel data
```{r}
panel_co2 <- pdata.frame(co2, index = c("country", "year"))
panel_co2 <- na.omit(panel_co2)

# All South Asia but without GFCF
sanog <- pdata.frame(sanog, index = c("country", "year"))
# Without Nepal but with all factors including GFCF
# sann <- pdata.frame(sann, index = c("country", "year"))
# # All South Asia but without GFCF and Fossil Fuel
# sanf <- pdata.frame(sanf, index = c("country", "year"))
# # All factors except Fossil fuel, no nepal
# sann_nof <- pdata.frame(sann_nof, index = c("country", "year"))
```
### Root Unit tests
#### ADF tests
```{r}
adf.test(sanog_differenced$ln_forest, k=2)
#"ln_agri", "lnGDP", "ln_energy", "lnCO2", "lnCO2_wdi"
#ln_population, ln_urban, ln_forest
```
The ADF test shows all variables are non-stationary, thus we first-order difference the variables to make them stationary
#### Im, Pesaran, and Shin (IPS) test
```{r}
purtest(lnCO2 ~ 1, data = sanog, test = "ips", model = "pooling", lags = 2)  # Change the number of lags as needed

# Perform the KPSS test for a specific variable
kpss.test(sanog$ln_forest, null = "Trend")

# Print the test result
cat("KPSS Test Statistic:", kpss_result$statistic, "\n")
cat("P-value:", kpss_result$p.value, "\n")
```

### Creating a differenced panel dataset
Following variables are stationary after one differencing: ln_agri, lnGDP, ln_energy, lnCO2, lnCO2_wdi.
ln_population, ln_urban, ln_forest are stationary after second order differencing

Dependent variables:lnCO2_wdi, lnCO2
Independent variables: ln_forest, ln_agri, lnGDP, ln_population, ln_urban, ln_energy
#### SANOG - All South Asia but without GFCF
```{r}
sanog_differenced_vars <- setdiff(names(sanog), c("year", "country"))
# panel_co2$Fossil.fuel.energy.consumption....of.total. <- NULL
# Perform first-order differencing for selected variables
sanog_differenced <- sanog %>%
  group_by(country) %>%  # Assuming you have a variable for country or entity
  arrange(country, year) %>%  # Arrange by country and time
  mutate(across(all_of(sanog_differenced_vars), ~ c(NA, diff(.)))) %>%
  ungroup()

sanog_differenced <- na.omit(sanog_differenced)

sanog_differenced_vars2 <- setdiff(names(sanog), c("year", "country", "ln_agri", "lnGDP", "lnGDPsq", "ln_fossil", "ln_energy", "lnCO2", "lnCO2_wdi"))

# Perform second-order differencing for selected variables
sanog_differenced2 <- sanog_differenced %>%
  group_by(country) %>%  # Assuming you have a variable for country or entity
  arrange(country, year) %>%  # Arrange by country and time
  mutate(across(all_of(sanog_differenced_vars2), ~ c(NA, diff(.)))) %>%
  ungroup()

sanog_differenced2 <- na.omit(sanog_differenced2)
sanog_panel <- pdata.frame(sanog_differenced2, index = c("country", "year"))
```

#### SANN - Without Nepal but with all factors including GFCF
```{r}
sann_differenced_vars <- setdiff(names(sann), c("year", "country"))
# panel_co2$Fossil.fuel.energy.consumption....of.total. <- NULL
# Perform first-order differencing for selected variables
sann_differenced <- sann %>%
  group_by(country) %>%  # Assuming you have a variable for country or entity
  arrange(country, year) %>%  # Arrange by country and time
  mutate(across(all_of(sann_differenced_vars), ~ c(NA, diff(.)))) %>%
  ungroup()

sann_differenced <- na.omit(sann_differenced)

sann_differenced_vars2 <- setdiff(names(sann), c("year", "country", "ln_agri", "lnGDP", "ln_fossil", "ln_energy", "lnCO2", "lnCO2_wdi"))

# Perform second-order differencing for selected variables
sann_differenced2 <- sann_differenced %>%
  group_by(country) %>%  # Assuming you have a variable for country or entity
  arrange(country, year) %>%  # Arrange by country and time
  mutate(across(all_of(sann_differenced_vars2), ~ c(NA, diff(.)))) %>%
  ungroup()

sann_differenced2 <- na.omit(sann_differenced2)
sann_panel <- pdata.frame(sann_differenced2, index = c("country", "year"))
```
#### SANF - All South Asia but without GFCF and Fossil Fuel
```{r}
sanf_differenced_vars <- setdiff(names(sanf), c("year", "country"))
# panel_co2$Fossil.fuel.energy.consumption....of.total. <- NULL
# Perform first-order differencing for selected variables
sanf_differenced <- sanf %>%
  group_by(country) %>%  # Assuming you have a variable for country or entity
  arrange(country, year) %>%  # Arrange by country and time
  mutate(across(all_of(sanf_differenced_vars), ~ c(NA, diff(.)))) %>%
  ungroup()

sanf_differenced <- na.omit(sanf_differenced)

sanf_differenced_vars2 <- setdiff(names(sanf), c("year", "country", "ln_agri", "lnGDP", "ln_energy", "lnCO2", "lnCO2_wdi"))

# Perform second-order differencing for selected variables
sanf_differenced2 <- sanf_differenced %>%
  group_by(country) %>%  # Assuming you have a variable for country or entity
  arrange(country, year) %>%  # Arrange by country and time
  mutate(across(all_of(sanf_differenced_vars2), ~ c(NA, diff(.)))) %>%
  ungroup()

sanf_differenced2 <- na.omit(sanf_differenced2)
sanf_panel <- pdata.frame(sanf_differenced2, index = c("country", "year"))
```
#### SANN_NOF - All factors except Fossil fuel, no nepal
```{r}
sann_nof_differenced_vars <- setdiff(names(sann_nof), c("year", "country"))
# panel_co2$Fossil.fuel.energy.consumption....of.total. <- NULL
# Perform first-order differencing for selected variables
sann_nof_differenced <- sann_nof %>%
  group_by(country) %>%  # Assuming you have a variable for country or entity
  arrange(country, year) %>%  # Arrange by country and time
  mutate(across(all_of(sann_nof_differenced_vars), ~ c(NA, diff(.)))) %>%
  ungroup()

sann_nof_differenced <- na.omit(sann_nof_differenced)

sann_nof_differenced_vars2 <- setdiff(names(sann_nof), c("year", "country", "ln_agri", "lnGDP", "ln_energy", "lnCO2", "lnCO2_wdi", "ln_capital"))

# Perform second-order differencing for selected variables
sann_nof_differenced2 <- sann_nof_differenced %>%
  group_by(country) %>%  # Assuming you have a variable for country or entity
  arrange(country, year) %>%  # Arrange by country and time
  mutate(across(all_of(sann_nof_differenced_vars2), ~ c(NA, diff(.)))) %>%
  ungroup()

sann_nof_differenced2 <- na.omit(sann_nof_differenced2)
sann_nof_panel <- pdata.frame(sann_nof_differenced2, index = c("country", "year"))
```

## Multicollinearity tests
### Checking multicollinearity - SANOG
```{r}
correlation_matrix_sanog <- cor(sanog_panel[, c("lnGDP", "ln_urban", "ln_population","ln_energy", "ln_fossil","ln_agri", "ln_forest")])
correlation_matrix_co2 <- cor(sanog_panel[, c("lnCO2", "lnCO2_wdi")])
high_correlations <- which(correlation_matrix_sanog > 0.5 & correlation_matrix_sanog < 0.7, arr.ind = TRUE)
very_high_correlations <- which(correlation_matrix_sanog > 0.7 & correlation_matrix_sanog < 1, arr.ind = TRUE)

# Extract variable pairs
variable_pairs <- data.frame(
  Variable1 = rownames(correlation_matrix_sanog)[high_correlations[, 1]],
  Variable2 = colnames(correlation_matrix_sanog)[high_correlations[, 2]]
  )
# Extract variable pairs
variable_pairs_high <- data.frame(
  Variable1 = rownames(correlation_matrix_sanog)[very_high_correlations[, 1]],
  Variable2 = colnames(correlation_matrix_sanog)[very_high_correlations[, 2]]
  )
# Examine the variable pairs
print(variable_pairs)
print(variable_pairs_high)
heatmap(correlation_matrix_sanog)

sanog_panel$GDP <- sanog_panel$lnGDP
sanog_panel$GDP_sq <- (sanog_panel$GDP)^2
sanog_panel$Urbanization <- sanog_panel$ln_urban
sanog_panel$Population <- sanog_panel$ln_population
sanog_panel$`Energy consumption` <- sanog_panel$ln_energy
sanog_panel$`Fossil fuel share` <- sanog_panel$ln_fossil
sanog_panel$`Agricultural production` <- sanog_panel$ln_agri
sanog_panel$`Forest area` <- sanog_panel$ln_forest
sanog_panel$`Energy use` <- sanog_panel$`Energy consumption`
sanog_panel$`Fossil fuel` <- sanog_panel$`Fossil fuel share`
sanog_panel$Forest <- sanog_panel$`Forest area`
sanog_panel$Agriculture <- sanog_panel$`Agricultural production`
sanog_panel$CO2 <- sanog_panel$lnCO2_wdi
sanog_panel$`CO2(GCB)` <- sanog_panel$lnCO2


correlation_matrix_sanog <- cor(sanog_panel[, c("GDP", "Urbanization", "Population", "Energy use" , "Fossil fuel", "Agriculture", "Forest")])
my_palette <- colorRampPalette(c("blue", "white", "red"))(50)

# Create the heatmap with custom settings

heatmap(
  correlation_matrix_sanog,
  main = "Correlation Heatmap",  # Main title
  margins = c(7, 7),          # Adjust margins
  cex.main = 0.8,               # Main title font size
  cex.axis = 0.8,               # Axis label font size
  cex.lab = 0.8,                # Variable labels font size
  symm = FALSE                   # Display a symmetric heatmap
)

colnames(sanog_panel)
correlation_matrix_sanog
```

### Checking multicollinearity - SANN
```{r}
correlation_matrix_sann <- cor(sann_panel[, c("ln_agri", "lnGDP", "ln_energy", "ln_population" , "ln_urban", "ln_forest", "ln_fossil", "ln_capital")])

heatmap(correlation_matrix_sann)
```
### Checking multicollinearity - SANF
```{r}
correlation_matrix_sanf <- cor(sanf_panel[, c("ln_agri", "lnGDP", "ln_energy", "ln_population" , "ln_urban", "ln_forest")])
heatmap(correlation_matrix_sanf)
```

### Checking multicollinearity - SANF
```{r}
correlation_matrix_sann_nof <- cor(sann_nof_panel[, c("ln_agri", "lnGDP", "ln_capital", "ln_population" , "ln_urban", "ln_forest")])
heatmap(correlation_matrix_sann_nof)
```


### Pooled OLS test 
```{r}
sanog

pooltest(lnCO2 ~ ln_agri + lnGDP + ln_energy + ln_population + ln_urban + ln_forest + ln_fossil, data = sanog_panel, model = "within")
## The pool test shows that the pooled data in unstable. We can't do pooled OLS regression
```
# Regressions
## SANOG
### Pooled effect - sanog
```{r}
pool_sanog <- plm(lnCO2_wdi ~ ln_agri + lnGDP + ln_energy + ln_population + ln_urban + ln_forest + ln_fossil,data = sanog_panel, model = "pooling")
summary(pool_sanog)
```

### Fixed effect - sanog
```{r}
fesanog <- plm(lnCO2_wdi ~ lnGDP + ln_urban + ln_population + ln_energy + ln_fossil + ln_agri + ln_forest,
               data = sanog_panel,model = "within")
summary(fesanog)
coeftest(fesanog, vcovHC(fesanog, method = "arellano"))

fesanog_gcb <- plm(lnCO2 ~ lnGDP + ln_urban + ln_population + ln_energy + ln_fossil + ln_agri + ln_forest,
               data = sanog_panel,model = "within")
coeftest(fesanog_gcb, vcovHC(fesanog_gcb, method = "arellano"))
```
### Random effect - sanog
```{r}
resanog <-  plm(lnCO2_wdi ~ lnGDP + ln_urban + ln_population + ln_energy + ln_fossil + ln_agri + ln_forest, data = sanog_panel, model = "random", random.method = "walhus")
coeftest(resanog, vcovHC(resanog, method = "arellano"))

resanog_gcb <- plm(lnCO2 ~ lnGDP + ln_urban + ln_population + ln_energy + ln_fossil + ln_agri + ln_forest,data = sanog_panel, model = "random", random.method = "walhus")
coeftest(resanog_gcb, vcovHC(resanog_gcb, method = "arellano"))
```

## SANN
### Fixed effect - sann
```{r}
fesann <- plm(lnCO2_wdi ~ ln_agri + lnGDP + ln_energy + ln_population + ln_urban + ln_forest + ln_fossil + ln_capital,
                           data = sann_panel,
                           model = "within"
                           )
summary(fesann)
```
### Random effect - sann
```{r}
resann <-  plm(lnCO2_wdi ~ ln_agri + lnGDP + ln_energy + ln_population + ln_urban + ln_forest + ln_fossil + ln_capital,
                 data = sann_panel, model = "random", random.method = "walhus")
summary(resann)
```

## SANF
### Fixed effect - sanf
```{r}
fesanf <- plm(lnCO2_wdi ~ ln_agri + lnGDP + ln_energy + ln_population + ln_urban + ln_forest,
                           data = sann_panel,model = "within")
summary(fesanf)
```
### Random effect - sanf
```{r}
resanf <-  plm(lnCO2_wdi ~ ln_agri + lnGDP + ln_energy + ln_population + ln_urban + ln_forest,
                 data = sann_panel, model = "random", random.method = "walhus")
summary(resanf)
```
## SANN_NOF
### Fixed effect - sann_nof
```{r}
fesann_nof <- plm(lnCO2_wdi ~ ln_agri + lnGDP + ln_energy + ln_population + ln_urban + ln_forest + ln_capital,
                           data = sann_nof_panel,model = "within")
summary(fesann_nof)
```
### Random effect - sann_nof
```{r}
resann_nof <-  plm(lnCO2_wdi ~ ln_agri + lnGDP + ln_energy + ln_population + ln_urban + ln_forest + ln_capital,
                 data = sann_nof_panel, model = "random", random.method = "walhus")
summary(resann_nof)
```

# Tests
### Pooled vs Fixed
```{r}
pooltest(lnCO2_wdi ~ ln_agri + lnGDP + ln_energy + ln_population + ln_urban + ln_forest + ln_fossil,
         data = sanog_panel, model = "within")

pFtest(fesanog, pool_sanog)
```
Pooled OLS model works in this case
### Test for individual and time effect in panel data
```{r}
plmtest(lnCO2_wdi ~ ln_agri + lnGDP + ln_energy + ln_population + ln_urban + ln_forest + ln_fossil,
        data = sanog_panel, effect = "twoways", type = "ghm")
```
No significant individual and time effects
### Hausman test
```{r}
phtest(fesanog, resanog)
```

### Autocorrelation test
```{r}
bgtest(resanog)
bgtest(resanf)
bgtest(resann)
bgtest(resann_nof)
```
### Heteroskasdicity test
```{r}
bptest(lnCO2 ~ lnGDP + ln_agri + ln_population + ln_urban + ln_energy_consumption,
       data = sanog_panel, studentize = F)
```
### Robust Regression coefficients
```{r}
fesanog_robust <- coeftest(fesanog, vcovHC(fesanog, method = "arellano"))
resanog_robust <- coeftest(resanog, vcovHC(resanog, method = "arellano"))

fesanog_robust_gcb <- coeftest(fesanog_gcb, vcovHC(fesanog_gcb, method = "arellano"))
resanog_robust_gcb <- coeftest(resanog_gcb, vcovHC(resanog_gcb, method = "arellano"))

# Saving standard errors
resanog_se <- coeftest(resanog, vcov.=vcovHC(resanog, method = c("arellano")))
fesanog_se <- coeftest(fesanog, vcov.=vcovHC(fesanog, method = c("arellano")))

resanog_gcb_se <- coeftest(resanog_gcb, vcov.=vcovHC(resanog_gcb, method = c("arellano")))
fesanog_gcb_se <- coeftest(fesanog_gcb, vcov.=vcovHC(fesanog_gcb, method = c("arellano")))
```

### Robust regression graph
```{r}
betas <- coef(resanog_robust)[-1]
cis <- confint(resanog_robust)

plot(betas, xlab = "" , main = "RE model (Data source: WDI)", ylab = "β Coefficients", xaxt = "n", pch = 1, ylim = c(-15, 15))
axis(1, at = c(1,2,3,4,5,6,7),
     labels = c("GDP", "Urban \n population", "Total \n population", "Energy \n consumption", "Fossil Fuel \n Share", "Agriculture \n Production", "Forest \n area"),
     cex.axis = 0.8)
lines(c(1,1), c(cis[2,1],cis[2,2]))
lines(c(2,2), c(cis[3,1],cis[3,2]))
lines(c(3,3), c(cis[4,1],cis[4,2]))
lines(c(4,4), c(cis[5,1],cis[5,2]))
lines(c(5,5), c(cis[6,1],cis[6,2]))
lines(c(6,6), c(cis[7,1],cis[7,2]))
lines(c(7,7), c(cis[8,1],cis[8,2]))
abline(h=0, lty = 2)

betas_fesanog_robust <- coef(fesanog_robust)[-1]
cis_fesanog_robust <- confint(fesanog_robust)

plot(betas, xlab = "" , main = "FE model (Data source: WDI)", ylab = "β Coefficients", xaxt = "n", pch = 1, ylim = c(-15, 15))
axis(1, at = c(1,2,3,4,5,6,7),
     labels = c("GDP", "Urban \n population", "Total \n population", "Energy \n consumption", "Fossil Fuel \n Share", "Agriculture \n Production", "Forest \n area"),
     cex.axis = 0.8)
lines(c(1,1), c(cis[2,1],cis[2,2]))
lines(c(2,2), c(cis[3,1],cis[3,2]))
lines(c(3,3), c(cis[4,1],cis[4,2]))
lines(c(4,4), c(cis[5,1],cis[5,2]))
lines(c(5,5), c(cis[6,1],cis[6,2]))
lines(c(6,6), c(cis[7,1],cis[7,2]))
lines(c(7,7), c(cis[8,1],cis[8,2]))
abline(h=0, lty = 2)

betas_fesanog_robust_gcb <- coef(fesanog_robust_gcb)[-1]
cis_fesanog_robust <- confint(fesanog_robust_gcb)

plot(betas, xlab = "" , main = "FE model (Data source: GCB)", ylab = "β Coefficients", xaxt = "n", pch = 1, ylim = c(-15, 15))
axis(1, at = c(1,2,3,4,5,6,7),
     labels = c("GDP", "Urban \n population", "Total \n population", "Energy \n consumption", "Fossil Fuel \n Share", "Agriculture \n Production", "Forest \n area"),
     cex.axis = 0.8)
lines(c(1,1), c(cis[2,1],cis[2,2]))
lines(c(2,2), c(cis[3,1],cis[3,2]))
lines(c(3,3), c(cis[4,1],cis[4,2]))
lines(c(4,4), c(cis[5,1],cis[5,2]))
lines(c(5,5), c(cis[6,1],cis[6,2]))
lines(c(6,6), c(cis[7,1],cis[7,2]))
lines(c(7,7), c(cis[8,1],cis[8,2]))
abline(h=0, lty = 2)

betas_resanog_robust_gcb <- coef(resanog_robust_gcb)[-1]
cis_resanog_robust <- confint(resanog_robust_gcb)

plot(betas, xlab = "" , main = "FE model (Data source: GCB)", ylab = "β Coefficients", xaxt = "n", pch = 1, ylim = c(-15, 15))
axis(1, at = c(1,2,3,4,5,6,7),
     labels = c("GDP", "Urban \n population", "Total \n population", "Energy \n consumption", "Fossil Fuel \n Share", "Agriculture \n Production", "Forest \n area"),
     cex.axis = 0.8)
lines(c(1,1), c(cis[2,1],cis[2,2]))
lines(c(2,2), c(cis[3,1],cis[3,2]))
lines(c(3,3), c(cis[4,1],cis[4,2]))
lines(c(4,4), c(cis[5,1],cis[5,2]))
lines(c(5,5), c(cis[6,1],cis[6,2]))
lines(c(6,6), c(cis[7,1],cis[7,2]))
lines(c(7,7), c(cis[8,1],cis[8,2]))
abline(h=0, lty = 2)

```

# LSDV model
```{r}
dummy <-lm(lnCO2_wdi ~ ln_agri + lnGDP + ln_energy + ln_population + ln_urban + ln_forest + ln_fossil +
                factor(country)-1, data = sanog_panel)
summary(dummy)
```
# Seemingly Unrelated Regression (SUR)
```{r}
sur_re <- systemfit(lnCO2_wdi ~ ln_agri + lnGDP + ln_energy + ln_population + ln_urban + ln_forest + ln_fossil,
                    method = "SUR", data = sanog_panel)
summary(sur_re)
```
# Print the regression table with robust std errors
```{r}
stargazer(resanog, type = "latex", title = "Regression Results", 
          align = TRUE, single.row = TRUE, se = list(vcovHC(resanog, type = "arellano")), out = "regression_table.html")

# Create labels for each model
model_labels <- c("Fixed effects", "Random effects")

resanog_gcb <- plm(lnCO2 ~ lnGDP + ln_urban + ln_population + ln_energy + ln_fossil + ln_agri + ln_forest,data = sanog_panel, model = "random", random.method = "walhus")

# Use stargazer to generate a table with three models side by side
stargazer(
  fesanog, resanog,
  type = "latex",
  title = "Regression results",
  align = TRUE,
  dep.var.labels=c("CO2 emissions"), covariate.labels=c("GDP","Urbanization", "Population","Energy Consumption","Fossil fuel", "Agriculture","Forest area"),
  omit.stat=c("LL","ser","f"),
  se = list((fesanog_se[,2]),
            (resanog_se[,2])),
  column.labels = model_labels,
  out = "regression_table.tex")


# Use stargazer to generate a table with three models side by side
stargazer(
  fesanog_gcb, resanog_gcb,
  type = "latex",
  title = "Regression results",
  align = TRUE,
  dep.var.labels=c("CO2 emissions"), covariate.labels=c("GDP","Urbanization", "Population","Energy Consumption","Fossil fuel", "Agriculture","Forest area"),
  omit.stat=c("LL","ser","f"),
  se = list((fesanog_se[,2]),
            (resanog_se[,2])),
  column.labels = model_labels,
  out = "regression_table_gcb.tex")

stargazer(correlation_matrix_sanog)
```

# Experiment 
```{r}
co2_exp <- co2 %>%
  select(country, year, `Fossil fuel energy consumption (% of total)`, `GDP (constant 2015 US$)`, `Population, total`, `Urban population (% of total population)`, `energy_consumption(TWh)`, `co2_emissions_our_world`, `Cereal production (metric tons)`)

co2_exp$lnCO2 <- log(co2_exp$`co2_emissions_our_world`)
co2_exp$ln_agri <- log(co2_exp$`Cereal production (metric tons)`)
co2_exp$lnGDP <- log(co2_exp$`GDP (constant 2015 US$)`)
co2_exp$ln_population <- log(co2_exp$`Population, total`)
co2_exp$ln_urban <- log(co2_exp$`Urban population (% of total population)`)
co2_exp$ln_energy <- log(co2_exp$`energy_consumption(TWh)`)
co2_exp$ln_fossil <- log(co2_exp$`Fossil fuel energy consumption (% of total)`)

co2_exp <- co2_exp %>%
  select(country, year, lnCO2, lnGDP, ln_population, ln_urban, ln_energy, ln_agri, ln_fossil)

co2_exp_differenced_vars <- setdiff(names(co2_exp), c("year", "country"))

# Perform first-order differencing for selected variables
co2_exp_dif <- co2_exp %>%
  group_by(country) %>%  # Assuming you have a variable for country or entity
  arrange(country, year) %>%  # Arrange by country and time
  mutate(across(all_of(co2_exp_differenced_vars), ~ c(NA, diff(.)))) %>%
  ungroup()

co2_exp_dif <- na.omit(co2_exp_dif)

co2_exp_differenced_vars2 

# Perform second-order differencing for selected variables
co2_exp_dif2 <- co2_exp_dif %>%
  group_by(country) %>%  # Assuming you have a variable for country or entity
  arrange(country, year) %>%  # Arrange by country and time
  mutate(across(all_of("ln_population"), ~ c(NA, diff(.)))) %>%
  ungroup()

co2_exp_dif2 <- na.omit(co2_exp_dif2)

co2_exp_dift1 <- co2_exp_dif2 %>%
  filter(year < 2001)

co2_exp_dift2 <- co2_exp_dif2 %>%
  filter(year > 2000)

co2_exp_panelt1 <- pdata.frame(co2_exp_dift1, index = c("country", "year"))
co2_exp_panelt2 <- pdata.frame(co2_exp_dift2, index = c("country", "year"))
co2_exp_panel <- pdata.frame(co2_exp_dif2, index = c("country", "year"))

co2_exp_ret1 <- plm(lnCO2 ~ ln_agri + lnGDP + ln_energy + ln_population + ln_urban + ln_fossil,
               data = co2_exp_panelt1, model = "random", random.method = "walhus")
coeftest(co2_exp_ret1, vcovHC(co2_exp_ret1, method = "arellano"))

co2_exp_fet1 <- plm(lnCO2 ~ ln_agri + lnGDP + ln_energy + ln_population + ln_urban + ln_fossil,
               data = co2_exp_panelt1, model = "within")
coeftest(co2_exp_fet1, vcovHC(co2_exp_fet1, method = "arellano"))

co2_exp_ret2 <- plm(lnCO2 ~ ln_agri + lnGDP + ln_energy + ln_population + ln_urban + ln_fossil,
               data = co2_exp_panelt2, model = "random", random.method = "walhus")
coeftest(co2_exp_ret2, vcovHC(co2_exp_ret2, method = "arellano"))

co2_exp_fet2 <- plm(lnCO2 ~ ln_agri + lnGDP + ln_energy + ln_population + ln_urban + ln_fossil,
               data = co2_exp_panelt2, model = "within")
coeftest(co2_exp_fet2, vcovHC(co2_exp_fet2, method = "arellano"))

co2_exp_fe <- plm(lnCO2 ~ ln_agri + lnGDP + ln_energy + ln_population + ln_urban + ln_fossil,
               data = co2_exp_panel, model = "within")
coeftest(co2_exp_fe, vcovHC(co2_exp_fe, method = "arellano"))

co2_exp_re <- plm(lnCO2 ~ ln_agri + lnGDP + ln_energy + ln_population + ln_urban + ln_fossil,
               data = co2_exp_panel, model = "random", random.method = "walhus")
coeftest(co2_exp_re, vcovHC(co2_exp_re, method = "arellano"))
```
# Experiment 2
```{r}
co2_exp <- co2 %>%
  select(country, year, `Fossil fuel energy consumption (% of total)`, `GDP (constant 2015 US$)`, `Population, total`, `Urban population (% of total population)`, `energy_consumption(TWh)`, `co2_emissions_our_world`, `Cereal production (metric tons)`)

co2_exp$lnCO2 <- log(co2_exp$`co2_emissions_our_world`)
co2_exp$ln_agri <- log(co2_exp$`Cereal production (metric tons)`)
co2_exp$lnGDP <- log(co2_exp$`GDP (constant 2015 US$)`)
co2_exp$ln_population <- log(co2_exp$`Population, total`)
co2_exp$ln_urban <- log(co2_exp$`Urban population (% of total population)`)
co2_exp$ln_energy <- log(co2_exp$`energy_consumption(TWh)`)
co2_exp$ln_fossil <- log(co2_exp$`Fossil fuel energy consumption (% of total)`)

co2_exp <- co2_exp %>%
  select(country, year, lnCO2, lnGDP, ln_population, ln_urban, ln_energy, ln_agri, ln_fossil)

co2_exp_differenced_vars <- setdiff(names(co2_exp), c("year", "country"))

# Perform first-order differencing for selected variables
co2_exp_dif <- co2_exp %>%
  group_by(country) %>%  # Assuming you have a variable for country or entity
  arrange(country, year) %>%  # Arrange by country and time
  mutate(across(all_of(co2_exp_differenced_vars), ~ c(NA, diff(.)))) %>%
  ungroup()

co2_exp_dif <- na.omit(co2_exp_dif)

co2_exp_differenced_vars2 

# Perform second-order differencing for selected variables
co2_exp_dif2 <- co2_exp_dif %>%
  group_by(country) %>%  # Assuming you have a variable for country or entity
  arrange(country, year) %>%  # Arrange by country and time
  mutate(across(all_of("ln_population"), ~ c(NA, diff(.)))) %>%
  ungroup()

co2_exp_dif2 <- na.omit(co2_exp_dif2)

co2_t2_panel <- pdata.frame(co2_t2, index = c("country", "year"))


re_t1_exp <- plm(lnCO2_wdi ~ ln_agri + lnGDP + ln_energy + ln_population + ln_urban + ln_forest + ln_fossil,
               data = co2_t1, model = "random", random.method = "walhus")

re_t2_exp <- plm(lnCO2_wdi ~ ln_agri + lnGDP + ln_energy + ln_population + ln_urban + ln_forest + ln_fossil,
               data = co2_t2, model = "random", random.method = "walhus")

coeftest(re_t1_exp, vcovHC(re_t1_exp, method = "arellano"))
coeftest(re_t2_exp, vcovHC(re_t2_exp, method = "arellano"))

re_t1_pool <- plm(lnCO2_wdi ~ ln_agri + lnGDP + ln_energy + ln_population + ln_urban + ln_forest + ln_fossil,
               data = co2_t1, model = "pooling")

re_t2_pool <- plm(lnCO2_wdi ~ ln_agri + lnGDP + ln_energy + ln_population + ln_urban + ln_forest + ln_fossil,
               data = co2_t2, model = "pooling")

coeftest(re_t1_pool, vcovHC(re_t1_pool, method = "arellano"))
coeftest(re_t2_pool, vcovHC(re_t2_pool, method = "arellano"))

fe_t1 <- plm(lnCO2_wdi ~ ln_agri + lnGDP + ln_energy + ln_population + ln_urban + ln_forest + ln_fossil,
               data = co2_t1, model = "within")

fe_t2 <- plm(lnCO2_wdi ~ ln_agri + lnGDP + ln_energy + ln_population + ln_urban + ln_forest + ln_fossil,
               data = co2_t2, model = "within")

coeftest(fe_t1, vcovHC(fe_t1, method = "arellano"))
coeftest(fe_t2, vcovHC(fe_t2, method = "arellano"))
```


# Descriptive statistics
```{r}
sanog_2 <- sanog %>% 
  group_by(country)

sanog_2$`CO2 emissions (million ton)`  <- (sanog_2$`CO2 emissions (kt)`)/(10^3)
sanog_2$`CO2 emissions (GCB)`  <- (sanog_2$co2_emissions_our_world)/(10^6)
sanog_2$`Population (mil)`  <- (sanog_2$`Population, total`)/(10^6)
sanog_2$`Agriculture production (mil metric tons)`  <- (sanog_2$`Cereal production (metric tons)`)/(10^6)
sanog_2$`GDP (trillion)`  <- sanog_2$`GDP (constant 2015 US$)`/(10^9)


table1::label(sanog_2$`energy_consumption(TWh)`) <- "Energy Consumption (TWh)"
table1::label(sanog_2$`GDP (trillion)`) <- "GDP (USD trillion)"


my_table <- table1::table1(~`CO2 emissions (million ton)` + `CO2 emissions (GCB)` + `GDP (trillion)` + `Urban population (% of total population)` + `Population (mil)` + `energy_consumption(TWh)` + `Fossil fuel energy consumption (% of total)` + `Agriculture production (mil metric tons)` + `Forest area (% of land area)` | country, data = sanog_2)

latex_code <- latex_table(my_table)

t1kable(my_table)
# Save the table to a CSV file
write.csv(my_table, file = "my_table.csv", row.names = FALSE)

# Or, save the table to a text file
write.table(my_table, file = "my_table.txt", sep = "\t", row.names = FALSE)

```


# Print the regression table with robust std errors
```{r}
stargazer(resanog, type = "latex", title = "Regression Results", 
          align = TRUE, single.row = TRUE, se = list(vcovHC(resanog, type = "arellano")), out = "regression_table.html")

# Create labels for each model
model_labels <- c("FE model", "RE model")

# Use stargazer to generate a table with three models side by side
stargazer(
  fesanog, resanog,
  type = "html",
  title = "Regression results",
  align = TRUE,
  single.row = TRUE,
  se = list((fesanog_se[,2]),
            (resanog_se[,2])),
  column.labels = model_labels,
  out = "regression_table.html")

# Use stargazer to generate a table with three models side by side
stargazer(
  fesanog_gcb, resanog_gcb,
  type = "html",
  title = "Regression results",
  align = TRUE,
  single.row = TRUE,
  se = list((fesanog_gcb_se[,2]),
            (resanog_gcb_se[,2])),
  column.labels = model_labels,
  out = "regression_table_gcb.html")

```
