---
title: "Thesis graphs"
author: "Hassan Aziz"
date: "2023-10-29"
output: html_document
---
## Loading Packages
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(gplots)
library(dplyr)
library(arsenal)
library(knitr) 
library(stargazer)
library(effects)
library(readxl)
library(reshape2)
library(plm)
library(glmnet)
library(scales)
library(lmtest)
library(urca)
library(forecast)
library(tseries)
library(systemfit)
library(writexl)
library(janitor)
library(sandwich)
library(broom)
library(WDI)
options(scipen = 999)  
```

## WDI data
### Load the dataset
```{r}
wdi <- read_excel("P_Data_Extract_From_World_Development_Indicators-9.xlsx")

```
### Clean the dataset
```{r}
### Rename variables
names(wdi)[names(wdi) == "Country Name"] <- "country"
names(wdi)[names(wdi) == "Country Code"] <- "country_code"
names(wdi)[names(wdi) == "Series Name"] <- "series"

# Drop irrelevant variables
wdi$`Series Code` <- NULL

# Remove everything after the first space from column names using sub()
colnames(wdi) <- sub(" .*", "", colnames(wdi))

# Define the columns you want to exclude from conversion
columns_to_exclude <- c("country", "country_code", "series")

# Get the names of columns to convert (all except those in columns_to_exclude)
columns_to_convert <- setdiff(names(wdi), columns_to_exclude)

# Convert all columns except the specified ones to numeric
wdi[columns_to_convert] <- lapply(wdi[columns_to_convert], as.numeric)

# Change data from wide to long
wdi_long <- wdi %>%
  pivot_longer(cols = -c(country, country_code, series), 
               names_to = "Year",
               values_to = "Value")

# remove missing values
wdi_long <- na.omit(wdi_long)

# Change data from long back to wide
wdi_wide <- wdi_long %>%
  spread(key = series, value = Value)

names(wdi_wide)[names(wdi_wide) == "Year"] <- "year"
wdi_wide$year <- as.numeric(wdi_wide$year)
```

## Our World Data
### Loading and cleaning the dataset
```{r}
primary_energy <-read.csv("primary-energy-cons.csv")
emissions <- read.csv("annual-co2-emissions-per-country.csv")
temp <- merge(primary_energy, emissions, by = c("Entity", "Code", "Year"), all = FALSE)
names(temp)[names(temp) == "Primary.energy.consumption..TWh."] <- "energy_consumption(TWh)"
names(temp)[names(temp) == "Annual.CO..emissions"] <- "co2_emissions_our_world"
names(temp)[names(temp) == "Entity"] <- "country"
names(temp)[names(temp) == "Code"] <- "country_code"
names(temp)[names(temp) == "Year"] <- "year"
```
### Merging the two datasets
```{r}
co2 <- merge(wdi_wide, temp, by = c("country", "country_code", "year"), all = FALSE)
co2$country_code <- NULL
```

### Scaling the data
```{r}
co2$year <- as.numeric(co2$year)
## Log transformations
co2$lnCO2_wdi <- log(co2$`CO2 emissions (kt)`)
co2$lnCO2 <- log(co2$co2_emissions_our_world)

co2$ln_forest <- log(co2$`Forest area (sq. km)`)
co2$ln_agri <- log(co2$`Cereal production (metric tons)`)
co2$lnGDP <- log(co2$`GDP (constant 2015 US$)`)
co2$ln_population <- log(co2$`Population, total`)
co2$ln_urban <- log(co2$`Urban population (% of total population)`)
co2$ln_energy <- log(co2$`energy_consumption(TWh)`)
co2$gross_cap_by_gdp <- ((co2$`Gross fixed capital formation (constant 2015 US$)`)/(co2$`GDP (constant 2015 US$)`))*100
co2$ln_capital <- log(co2$gross_cap_by_gdp)
co2$ln_fossil <- log(co2$`Fossil fuel energy consumption (% of total)`)
co2$lnGDPsq <- log((co2$`GDP (constant 2015 US$)`)^2)
```


## Descriptive graphs
#### CO2
```{r}
Co2_graph <- ggplot(data = co2, aes(x = year, y = lnCO2, color = country, group = country)) +
  geom_line() +
  labs(x = "Year", y = "log of CO2 emissions (trillion tons)", color = "Country") +
  scale_color_manual(values = c("Pakistan" = "blue", "India" = "red", "Bangladesh" = "green", "Nepal" = "purple", "Sri Lanka" = "orange")) +
  scale_x_continuous(breaks = c(1975, 1985, 1995, 2005, 2015, 2020)) +
  theme_minimal() +
  theme(legend.position = "right") 

Co2_wdi_graph <- ggplot(data = sanf, aes(x = year, y = lnCO2_wdi, color = country, group = country)) +
  geom_line() +
  labs(x = "Year", y = "log of CO2 emissions (trillion tons)", color = "Country") +
  scale_color_manual(values = c("Pakistan" = "blue", "India" = "red", "Bangladesh" = "green", "Nepal" = "purple", "Sri Lanka" = "orange")) +
  scale_x_continuous(breaks = c(1990, 1995, 2000, 2005, 2010, 2015, 2020)) +
  theme_minimal() +
  theme(legend.position = "right") 

co2 <- co2 %>%
  filter(year > 1989 & year <2021)

ggplot(co2, aes(x = year)) +
  geom_line(aes(y = lnCO2_wdi, color = country, linetype = "lnCO2_wdi"), linewidth = 0.5) +
  geom_line(aes(y = lnGDP, color = country, linetype = "lnGDP"), linewidth = 0.5) +
  scale_color_manual(values = c("Pakistan" = "blue", "India" = "red", "Bangladesh" = "green", "Nepal" = "purple", "Sri Lanka" = "orange")) +
  scale_y_continuous("log of CO2 emissions", sec.axis = sec_axis(~., name = "log of GDP")) +
  scale_linetype_manual(values = c("lnCO2_wdi" = "solid", "lnGDP" = "dashed")) +
  labs(title = "Graph of CO2 emissions and GDP in South Asia",
       x = "Year",
       y = "lnCO2_wdi",
       color = "Country",
       linetype = "Linetype") +
  theme_minimal()

ggplot(sanf, aes(x = year)) +
  geom_line(aes(y = lnCO2_wdi, color = country, linetype = "lnCO2_wdi"), linewidth = 0.5) +
  geom_line(aes(y = ln_urban, color = country, linetype = "ln_urban"), linewidth = 0.5) +
  scale_color_manual(values = c("Pakistan" = "blue", "India" = "red", "Bangladesh" = "green", "Nepal" = "purple", "Sri Lanka" = "orange")) +
  scale_y_continuous("log of CO2 emissions", sec.axis = sec_axis(~., name = "log of urban population share")) +
  scale_linetype_manual(values = c("lnCO2_wdi" = "solid", "ln_urban" = "dashed")) +
  labs(title = "Graph of CO2 emissions and urban share in South Asia",
       x = "Year",
       y = "lnCO2_wdi",
       color = "Country",
       linetype = "Linetype") +
  theme_minimal()

Co2_wdi_graph

```
#### Population
```{r}
ln_population_graph <- ggplot(data = co2, aes(x = year, y = ln_population, color = country, group = country)) +
  geom_line() +
  labs(x = "Year", y = "log of population ", color = "Country") +
  scale_color_manual(values = c("Pakistan" = "blue", "India" = "red", "Bangladesh" = "green", "Nepal" = "purple", "Sri Lanka" = "orange")) +
  scale_x_continuous(breaks = c(1965, 1975, 1985, 1995, 2005, 2015, 2025)) +
  theme_minimal() +
  theme(legend.position = "right")

data <- WDI(country = "all", indicator = "SP.POP.TOTL", start = 1975, end = 2020)

named_observations <- c("World", "South Asia", "East Asia & Pacific")

subset_data <- data %>%
  filter(country %in% named_observations)

# Filter the dataset to include only "South Asia" observations
south_asia_data <- subset_data %>%
  filter(country == "South Asia")

# Filter the dataset to include only "World" observations
world_data <- subset_data %>%
  filter(country == "World")

# Calculate the population ratio for South Asia relative to the world population
result <- left_join(south_asia_data, world_data, by = "year") %>%
  mutate(Population_Ratio = SP.POP.TOTL.x / SP.POP.TOTL.y) %>%
  select(year, Population_Ratio)

### population ratio graph
ln_population_graph <- ggplot(data = result, aes(x = year, y = Population_Ratio)) +
  geom_line() +
  labs(x = "Year", y = "Population ratio (South Asia / World)", color = "Country") +
  scale_x_continuous(breaks = c(1970, 1980, 1990, 2000, 2010, 2020)) +
  theme_minimal() +
  theme(legend.position = "right")

### population growth graph
named_observations <- c("World", "South Asia", "East Asia & Pacific")

subset_data <- data %>%
  filter(country %in% named_observations)

subset_data <- subset_data %>%
  arrange(country, year)

# Calculate the growth rate
growth_rate <- c(NA, diff(subset_data$SP.POP.TOTL) / head(subset_data$SP.POP.TOTL, -1))

# Add the growth rate as a new column in the dataset
subset_data$GrowthRate <- (growth_rate*100)

subset_data <- subset_data %>%
  filter(year > 1975)

ggplot(data = subset_data, aes(x = year, y = GrowthRate, color = country, group = country)) +
  geom_line() +
  labs(x = "Year", y = "Population growth ", color = "Country") +
  scale_color_manual(values = c("East Asia & Pacific" = "blue", "South Asia" = "red", "World" = "black")) +
  scale_x_continuous(breaks = c(1975, 1985, 1995, 2005, 2015, 2025)) +
  theme_minimal() +
  theme(legend.position = "right")

### CO2 emissions per capita

data <- WDI(country = "all", indicator = "EN.ATM.CO2E.PC", start = 1990, end = 2020)

subset_data <- data %>%
  filter(country %in% named_observations)  

ggplot(data = subset_data, aes(x = year, y = EN.ATM.CO2E.PC, color = country, group = country)) +
  geom_line() +
  labs(x = "Year", y = "CO2 emissions per capita ", color = "Country") +
  scale_color_manual(values = c("East Asia & Pacific" = "black", "South Asia" = "pink", "World" = "grey")) +
  scale_x_continuous(breaks = c(1990, 1995, 2000, 2005, 2010, 2015, 2020)) +
  theme_minimal() +
  theme(legend.position = "right")
```
#### Urban Population
```{r}
ln_urban_graph <- ggplot(data = co2, aes(x = year, y = ln_urban, color = country, group = country)) +
  geom_line() +
  labs(x = "Year", y = "log of urban population", color = "Country") +
  scale_color_manual(values = c("Pakistan" = "blue", "India" = "red", "Bangladesh" = "green", "Nepal" = "purple", "Sri Lanka" = "orange")) +
  scale_x_continuous(breaks = c(1965, 1975, 1985, 1995, 2005, 2015, 2025)) +
  theme_minimal() +
  theme(legend.position = "right")

ln_urban_graph <- ggplot(data = sanf, aes(x = year, y = ln_urban, color = country, group = country)) +
  geom_line() +
  labs(x = "Year", y = "Log of urban population share", color = "Country") +
  scale_color_manual(values = c("Pakistan" = "blue", "India" = "red", "Bangladesh" = "green", "Nepal" = "purple", "Sri Lanka" = "orange")) +
  scale_x_continuous(breaks = c(1990, 1995, 2000, 2005, 2010, 2015, 2020)) +
  theme_minimal() +
  theme(legend.position = "right")
```
#### Manufacturing
```{r}
ln_manufacturing_graph <- ggplot(data = co2, aes(x = year, y = ln_manufacturing, color = country, group = country)) +
  geom_line() +
  labs(x = "Year", y = "log of manufacturing value (bn)", color = "Country") +
  scale_color_manual(values = c("Pakistan" = "blue", "India" = "red", "Bangladesh" = "green", "Nepal" = "purple", "Sri Lanka" = "orange")) +
  scale_x_continuous(breaks = c(1965, 1975, 1985, 1995, 2005, 2015, 2025)) +
  theme_minimal() +
  theme(legend.position = "right")
```
#### GDP
```{r}
ln_gdp_graph <- ggplot(data = co2, aes(x = year, y = lnGDP, color = country, group = country)) +
  geom_line() +
  labs(x = "Year", y = "log of GDP (bn)", color = "Country") +
  scale_color_manual(values = c("Pakistan" = "blue", "India" = "red", "Bangladesh" = "green", "Nepal" = "purple", "Sri Lanka" = "orange")) +
  scale_x_continuous(breaks = c(1965, 1975, 1985, 1995, 2005, 2015, 2025)) +
  theme_minimal() +
  theme(legend.position = "right")


```
#### Energy consumption
```{r}
ln_energy_graph <- ggplot(data = co2, aes(x = year, y = ln_energy, color = country, group = country)) +
  geom_line() +
  labs(x = "Year", y = "log of energy consumption, total", color = "Country") +
  scale_color_manual(values = c("Pakistan" = "blue", "India" = "red", "Bangladesh" = "green", "Nepal" = "purple", "Sri Lanka" = "orange")) +
  scale_x_continuous(breaks = c(1975, 1985, 1995, 2005, 2015)) +
  theme_minimal() +
  theme(legend.position = "right")

```
#### Fossil fuel 
```{r}
data <- WDI(country = "all", indicator = "EG.USE.COMM.FO.ZS", start = 1975, end = 2014)
named_observations <- c("India", "Pakistan", "Bangladesh", "Nepal", "Sri Lanka", "East Asia & Pacific", "World", "South Asia")

unique(data$country)
# Create a subset with only named observations using dplyr
subset_data <- data %>%
  filter(country %in% named_observations)

ln_fossil_graph <- ggplot(data = subset_data, aes(x = year, y = EG.USE.COMM.FO.ZS, color = country, group = country)) +
  geom_line() +
  labs(x = "Year", y = "Share of fossil fuel in energy consumption", color = "Country") +
  scale_color_manual(values = c("South Asia" = "Violet", "World" = "grey", "East Asia & Pacific" = "black", "Pakistan" = "blue", "India" = "red", "Bangladesh" = "green", "Nepal" = "purple", "Sri Lanka" = "orange")) +
  scale_x_continuous(breaks = c(1975, 1985, 1995, 2005, 2015, 2025)) +
  theme_minimal() +
  theme(legend.position = "right")
```

#### Agriculture
```{r}
data_agri <- WDI(country = "all", indicator = "AG.PRD.CREL.MT", start = 1975, end = 2020)
named_observations <- c("East Asia & Pacific", "World", "South Asia")
unique(data_agri$country)

subset_data_agri <- data_agri %>%
  filter(country %in% named_observations)

subset_data_agri <- subset_data_agri %>%
  arrange(country, year)

# Calculate the growth rate
growth_rate <- c(NA, diff(subset_data_agri$AG.PRD.CREL.MT) / head(subset_data_agri$AG.PRD.CREL.MT, -1))
subset_data_agri$GrowthRate <- (growth_rate*100)
subset_data_agri <- subset_data_agri %>%
  filter(year > 1975)

subset_data_agri <- subset_data_agri %>%
  group_by(country) %>%
  arrange(country,year) %>%
  mutate(
    Cumulative_Growth_Rate = cumsum(GrowthRate)
  ) %>%
  ungroup()

ln_agri <- ggplot(data = subset_data_agri, aes(x = year, y = Cumulative_Growth_Rate, color = country, group = country)) +
  geom_line() +
  labs(x = "Year", y = "Cereal production (cumulative growth rate)", color = "Country") +
  scale_color_manual(values = c("South Asia" = "Violet", "World" = "grey", "East Asia & Pacific" = "black")) +
  scale_x_continuous(breaks = c(1975, 1985, 1995, 2005, 2015, 2025)) +
  theme_minimal() +
  theme(legend.position = "right")

south_asia <- c("India", "Pakistan", "Bangladesh", "Nepal", "Sri Lanka")

sa_data_agri <- data_agri %>%
  filter(country %in% south_asia)

sa_agri <- ggplot(data = sa_data_agri, aes(x = year, y = AG.PRD.CREL.MT, color = country, group = country)) +
  geom_line() +
  labs(x = "Year", y = "Cereal production (cumulative growth rate)", color = "Country") +
  scale_color_manual(values = c("Pakistan" = "blue", "India" = "red", "Bangladesh" = "green", "Nepal" = "purple", "Sri Lanka" = "orange")) +
  scale_x_continuous(breaks = c(1975, 1985, 1995, 2005, 2015, 2025)) +
  theme_minimal() +
  theme(legend.position = "right")
```

##### Agri NO2 emissions
```{r}
data_agri_emissions <- WDI(country = "all", indicator = "EN.ATM.CO2E.PC", start = 1990, end = 2020)

subset_data_agri_emissions <- data_agri_emissions %>%
  filter(country %in% named_observations)

subset_data_agri_emissions <- subset_data_agri_emissions %>%
  arrange(country, year)

emissions_no2 <- ggplot(data = subset_data_agri_emissions, aes(x = year, y = EN.ATM.CO2E.PC, color = country, group = country)) +
  geom_line() +
  labs(x = "Year", y = "Agricultural NO2 emissions", color = "Country") +
  scale_color_manual(values = c("South Asia" = "Violet", "World" = "grey", "East Asia & Pacific" = "black")) +
  scale_x_continuous(breaks = c(1990, 2000, 2010, 2020)) +
  theme_minimal() +
  theme(legend.position = "right")

emissions_no2

```

##### Agri Methane emissions
```{r}
methane_emissions <- WDI(country = "all", indicator = "EN.ATM.METH.AG.ZS", start = 1990, end = 2020)

subset_methane_emissions <- methane_emissions %>%
  filter(country %in% named_observations)

subset_methane_emissions <- subset_methane_emissions %>%
  arrange(country, year)

emissions_methane <- ggplot(data = subset_methane_emissions, aes(x = year, y = EN.ATM.METH.AG.ZS, color = country, group = country)) +
  geom_line() +
  labs(x = "Year", y = "Agricultural NO2 emissions", color = "Country") +
  scale_color_manual(values = c("South Asia" = "Violet", "World" = "grey", "East Asia & Pacific" = "black")) +
  scale_x_continuous(breaks = c(1990, 2000, 2010, 2020)) +
  theme_minimal() +
  theme(legend.position = "right")

emissions_methane
```

##### Fertilizer use
```{r}
fert <- WDI(country = "all", indicator = "AG.CON.FERT.ZS", start = 1990, end = 2020)

subset_fert <- fert %>%
  filter(country %in% named_observations)

subset_fert <- subset_fert %>%
  arrange(country, year)

fert <- ggplot(data = subset_fert, aes(x = year, y = AG.CON.FERT.ZS, color = country, group = country)) +
  geom_line() +
  labs(x = "Year", y = "Fertilizer consumption (kg per hectare)", color = "Country") +
  scale_color_manual(values = c("South Asia" = "Violet", "World" = "grey", "East Asia & Pacific" = "black")) +
  scale_x_continuous(breaks = c(1990, 2000, 2010, 2020)) +
  theme_minimal() +
  theme(legend.position = "right")

fert
```


#### Forest
```{r}
wdi <- read_excel("P_Data_Extract_From_World_Development_Indicators-9.xlsx")
forest <- WDI(country = "all", indicator = "AG.LND.FRST.ZS", start = 1990, end = 2020)

forest <- forest %>%
  filter(country %in% south_asia)

ggplot(data = forest, aes(x = year, y = AG.LND.FRST.ZS, color = country, group = country)) +
  geom_line() +
  labs(x = "Year", y = "Forest area (% of land area)", color = "Country") +
  scale_color_manual(values = c("Pakistan" = "blue", "India" = "red", "Bangladesh" = "green", "Nepal" = "purple", "Sri Lanka" = "orange")) +
  scale_x_continuous(breaks = c(1990, 1995, 2000, 2005, 2010, 2015, 2020)) +
  theme_minimal() +
  theme(legend.position = "right")
```


#### Carbon intensity
```{r}
named_observations <- c("East Asia & Pacific", "World", "South Asia")
intensity <- WDI(country = "all", indicator = "EN.ATM.CO2E.PP.GD", start = 1990, end = 2020)

intensity <- intensity %>%
  filter(country %in% named_observations)

carbon_intensity_graph <- ggplot(data = intensity, aes(x = year, y = EN.ATM.CO2E.PP.GD, color = country, group = country)) +
  geom_line() +
  labs(x = "Year", y = "CO2 emissions (kg per $)", color = "Country") +
  scale_color_manual(values = c("East Asia & Pacific" = "black", "World" = "grey", "South Asia" = "pink")) +
  scale_x_continuous(breaks = c(1990, 2000, 2010, 2020)) +
  theme_minimal() +
  theme(legend.position = "right")
```
#### CO2wdi vs CO2(GCB)
```{r}
co2_gra <- co2 %>%
  filter(year > 1989 & year <2015)

ggplot(co2_gra, aes(x = year)) +
  geom_line(aes(y = lnCO2_wdi, color = country, linetype = "lnCO2_wdi"), linewidth = 0.5) +
  geom_line(aes(y = lnCO2, color = country, linetype = "lnCO2"), linewidth = 0.5) +
  scale_color_manual(values = c("Pakistan" = "blue", "India" = "red", "Bangladesh" = "green", "Nepal" = "purple", "Sri Lanka" = "orange")) +
  scale_y_continuous("log of CO2 emissions", sec.axis = sec_axis(~., name = "log of CO2 emissions (GCB)")) +
  scale_linetype_manual(values = c("lnCO2_wdi" = "solid", "lnCO2" = "dashed")) +
  labs(title = "Graph of CO2 emissions and GDP in South Asia",
       x = "Year",
       y = "lnCO2_wdi",
       color = "Country",
       linetype = "Linetype") +
  theme_bw()
```

